Modularizing Street Map Generation \\ \vspace{8pt} \large An Adventure in Tensors

### 2013-12-12 paper/paper.tex
fixed some references and typos in paper

The major eigenvectors of $t$ are
\[
    \left\{
        \lambda\mkvec{\cos{\theta}}{\sin{\theta}} ~|~ \lambda \neq 0
    \right\}
\]
The minor eigenvectors are
\[
    \left\{
        \lambda\mkvec
                {\cos{(\theta+\frac{\pi}{2})}}
                {\sin{(\theta+\frac{\pi}{2})}}
        ~|~ \lambda \neq 0
    \right\}
\]
A tensor field $T$ describes the flow of the constraints across a field,
and is represented as a function that maps from 2D points to tensors. In the
context of this project, a tensor is simply a tool we use to determine the
direction of a field at a point. This direction is found by calculating the
eigenvectors of the tensors. Major eigenvectors are parallel to the field;
minor eigenvectors are perpendicular.

@@

\begin{figure*}[t!]
\todo{reimplement pipeline image in \LaTeX}
\begin{center}
    \includegraphics[width=5.5in]{images/pipeline.pdf}
\end{center}
\caption{Overview of our pipeline}
\label{fig:pipeline}
\end{figure*}

@@

@@

\paragraph{Radial Constraints}
A radial constraint is used to describe specifically circular areas, such as
roundabouts or more generally curved roads, common in suburban areas. They are
defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric
circles
around the point, while minor ones emanate outwards. The basis field for a
radial constraint at a point $\pt=x_p,y_p$ is defined as
\[
    T(\pt) = \sqmat{y^2-x^2}{-2xy}{-2xy}{-(y^2-x^2)}
\]
where $x=x_p-x_0$ and $y=x_p-x_0$.

@@

\subsection{Streamline Tracing}
Mathematically, the basic streamline tracing algorithm is fairly simple,
however, in order to produce a set of streamlines that approximate a road map,
there are several important cases to consider. In order to trace the
streamlines, we are forced to use a heuristic, since it isn’t possible to trace
a continuous line over a constantly-changing tensor field. In order to
approximate this, we choose a step size $d_\textrm{step}$. Beginning at a seed
point, which is chosen using the methods detailed in section XX, we calculate
the eigenvector at that point, extend a line in that direction with length
$d_\textrm{step}$, and check if we have reached an end condition.  This process
is repeated until the line until one of three end conditions is reached:
\begin{enumerate}
    \item The point is outside the bounds of the map area
    \item The point is within a certain distance of an existing streamline.
    \item The point is part of a cycle
\end{enumerate}

@@

\begin{figure*}[t!]
    \todo{image of that nice circular constraint that worked really well with this}
\caption{Haskell implementation of eigenvector calculation}
% \label{fig:cycles} name?
\end{figure*}

@@

\subsection{Seed Placement}
To begin constructing a streamline, one must choose a seed point—an initial
coordinate in 2D space from which the streamline is traced. In practice, the
matter of carefully choosing seed points is very important in generating an
evenly spaced configuration of roads. We investigated three methods of placing
seed points:
\begin{enumerate}
    \item Random
    \item Furthest-Point
    \item Even Streamline Spacing
\end{enumerate}
Purely random placement of seed points generally results in both clusters of
roads which are too close together and large gaps between roads. In the
furthest-point placement, we place each new seed point within the bounds of the
space such that the distance to all existing streamlines is maximized. As a
simplifying heuristic for this method, we generate a number of random candidate
seed points which is small compared to the infinite number of possible seed
points (constrained only by floating point precision) and choose the one
furthest from all existing streamlines. In practice, this offers a significant
improvement over the random seed placement, and results in a more even
distribution of streamlines.

@@

-- returns a tuple of the (major, minor) eigenvectors for a tensor
eigenvectors :: Tensor -> (V2.Vector2, V2.Vector2)
eigenvectors (Tensor a 0.0 d) =
  if a > d then (V2.Vector2 1 0, V2.Vector2 0 1)
     else (V2.Vector2 0 1, V2.Vector2 1 0)
     eigenvectors (Tensor a b d) =
       let (eval1, eval2) = eigenvalues (Tensor a b d)
           mkEvec eval    = V2.Vector2 1 ((eval - a) / b)
       in (mkEvec eval1, mkEvec eval2)
\end{verbatim}
\caption{Haskell implementation of eigenvector calculation}
\label{fig:evecs}
\end{figure*}

\subsection{Modularity \& Extensibility}
We believe that we were successful in creating a modular system with reusable
parts that is easy to understand, extend, and modify. One way to illustrate
this is by looking at our pipeline. As a refresher, look at Fig. ***. Each
step of the pipeline has a clear correspondence in the modules that we have
written. We have two modules to deal with constraints, \texttt{Constraint}. To
transform
constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor}
modules, the latter
of which performs operations on tensors including eigenvector calculations, as
well as \texttt{Streamline}, which has the ability to calculate seed points
and trace
lines.

>>>>

The major eigenvectors of $t$ are
\[
    \left\{
        \lambda\mkvec{\cos{\theta}}{\sin{\theta}} ~|~ \lambda \neq 0
    \right\}
\]

The minor eigenvectors are
\[
    \left\{
        \lambda\mkvec
                {\cos{(\theta+\frac{\pi}{2})}}
                {\sin{(\theta+\frac{\pi}{2})}}
        ~|~ \lambda \neq 0
    \right\}
\]
A tensor field $T$ describes the flow of the constraints across a field,
and is represented as a function that maps from 2D points to tensors. In the
context of this project, a tensor is simply a tool we use to determine the
direction of a field at a point. This direction is found by calculating the
eigenvectors of the tensors. Major eigenvectors are parallel to the field;
minor eigenvectors are perpendicular.

@@

\begin{figure*}[t!]
\begin{center}
    \includegraphics[width=5.5in]{images/pipeline.pdf}
\end{center}
\caption{Overview of our pipeline}
\label{fig:pipeline}
\end{figure*}

@@

% \floatstyle{plain}
% \restylefloat{figure}
\begin{figure*}[t!]
    \centering
    \begin{subfigure}[b]{2.4in}
        \begin{verbatim}
            [{
              "tycon" : "Linear",
              "posx" : 2,
              "posy" : 2,
              "dir" : 0.01,
              "mag" : 3
            },
            {
              "tycon" : "Radial",
              "posx" : 10,
              "posy" : 10,
              "dir" : 0,
              "mag" : 0
            }]
        \end{verbatim}
        \caption{JSON file from by user}
    \end{subfigure}
    \hspace{20pt}
    \begin{subfigure}[b]{2.4in}
        \todo{image from constraints to left}
        \caption{generated constraints}
    \end{subfigure}
\caption{JSON parsing of constrains}
\label{fig:json}
\end{figure*}
% \floatstyle{boxed}
% \restylefloat{figure}

@@

\paragraph{Radial Constraints}
A radial constraint is used to describe specifically circular areas, such as
roundabouts or more generally curved roads, common in suburban areas. They are
defined at a point $(x_0,y_0)$; major streamlines are drawn as concentric
circles
around the point, while minor ones emanate outwards. The basis field for a
radial constraint at a point $\pt=x_p,y_p$ is defined as
\[
    T(\pt) = \sqmat{y^2-x^2}{-2xy}{-2xy}{-(y^2-x^2)}
\]
where $x=x_p-x_0$ and $y=y_p-y_0$.

@@

\subsection{Streamline Tracing}
Mathematically, the basic streamline tracing algorithm is fairly simple,
however, in order to produce a set of streamlines that approximate a road map,
there are several important cases to consider. In order to trace the
streamlines, we are forced to use a heuristic, since it isn’t possible to trace
a continuous line over a constantly-changing tensor field. In order to
approximate this, we choose a step size $d_\textrm{step}$. Beginning at a seed
point, which is chosen using the methods detailed in Section \ref{seedpoints},
we calculate the eigenvector at that point, extend a line in that direction
with length $d_\textrm{step}$, and check if we have reached an end condition.
This process is repeated until the line until one of three end conditions is
reached:
\begin{enumerate}
    \item The point is outside the bounds of the map area
    \item The point is within a certain distance of an existing streamline.
    \item The point is part of a cycle
\end{enumerate}

@@

\begin{figure*}[t!]
    \todo{image of that nice circular constraint that worked really well with
    this}
% \label{fig:cycles} name?
\end{figure*}

@@

\subsection{Seed Placement}\label{seedpoints}
To begin constructing a streamline, one must choose a seed point—an initial
coordinate in 2D space from which the streamline is traced. In practice, the
matter of carefully choosing seed points is very important in generating an
evenly spaced configuration of roads. We investigated three methods of placing
seed points:
\begin{enumerate}
    \item Random
    \item Furthest-Point
    \item Even Streamline Spacing
\end{enumerate}
Purely random placement of seed points generally results in both clusters of
roads which are too close together and large gaps between roads. In the
furthest-point placement, we place each new seed point within the bounds of the
space such that the distance to all existing streamlines is maximized. As a
simplifying heuristic for this method, we generate a number of random candidate
seed points which is small compared to the infinite number of possible seed
points (constrained only by floating point precision) and choose the one
furthest from all existing streamlines. In practice, this offers a significant
improvement over the random seed placement, and results in a more even
distribution of streamlines.

@@

-- returns a tuple of the (major, minor) eigenvectors for a tensor
eigenvectors :: Tensor -> (V2.Vector2, V2.Vector2)
eigenvectors (Tensor a 0.0 d) =
  if a > d then (V2.Vector2 1 0, V2.Vector2 0 1)
     else (V2.Vector2 0 1, V2.Vector2 1 0)
     eigenvectors (Tensor a b d) =
       let (eval1, eval2) = eigenvalues (Tensor a b d)
           mkEvec eval    = V2.Vector2 1 ((eval - a) / b)
       in (mkEvec eval1, mkEvec eval2)
\end{verbatim}
\caption{Haskell implementation of eigenvector calculation (from
            \texttt{Tensor.hs})}
\label{fig:evecs}
\end{figure*}

\subsection{Modularity \& Extensibility}
We believe that we were successful in creating a modular system with reusable
parts that is easy to understand, extend, and modify. One way to illustrate
this is by looking at our pipeline (Figure \ref{fig:pipeline}). Each step of
the pipeline has a clear correspondence in the modules that we have written.
We have two modules to deal with constraints, \texttt{Constraint}. To transform
constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor}
modules, the latter of which performs operations on tensors including
eigenvector calculations, as well as \texttt{Streamline}, which has the ability
to calculate seed points and trace lines.
